Structure Learning for $VAR(p)$ Models

### 2019-03-01 paper.tex
Some outlining and rough statements of theorems.

@@

\section{Theory}
\label{sec:theory}
Consider a wide sense stationary, vector valued and discrete time
stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
valued covariance sequence $R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ and spectra $S(\omega)$:

@@

\begin{definition}[Granger Causality]
  For the WSS series $x(t)$ we say that component $x_j$ \textit{Granger-Causes} (GC) component $x_i$ (with respect to $x(t)$) and write $x_j \rightarrow x_i$ if given Hilbert spaces $\mathcal{H}^{(i)}_{t - 1}$, $\mathcal{H}^{(i, j)}_{t - 1}$ generated by $\{x_i(t - \tau)\}_{\tau = 1}^\infty$ and $\{x_i(t - \tau), x_j(t - s\}_{\tau, s = 1}^\infty$ respectively we have

@@

\section{Structure Learning}
\subsection{Edge-Wise Grouped LASSO}
\subsection{Bayesian Posterior Thresholding}
\subsection{Pairwise Minimum Error Spanning Trees}
\section{Application}
\section{Conclusion}

>>>>

\def\gc{\overset{\text{GC}}{\rightarrow}}
\def\te{\overset{\mathcal{T}}{\rightarrow}}

@@

\section{Theory}
\label{sec:theory}
Consider a zero mean wide sense stationary, vector valued and discrete
time stochastic process $x(t) \in L^n_2(\Omega);\ t \in \Z$, having the matrix
valued covariance sequence
$R(\tau) \overset{\Delta}{=} \E x(t)x(t - \tau)^\T$ and spectra $S(\omega)$:

@@

\begin{definition}[Granger Causality]
  For the WSS series $x(t)$ we say that component $x_j$
  \textit{Granger-Causes} (GC) component $x_i$ (with respect to
  $x(t)$) and write $x_j \gc x_i$ if given Hilbert spaces
  $\mathcal{H}^{(i)}_{t - 1}$, $\mathcal{H}^{(i, j)}_{t - 1}$
  generated by $\{x_i(t - \tau)\}_{\tau = 1}^\infty$ and
  $\{x_i(t - \tau), x_j(t - s\}_{\tau, s = 1}^\infty$ respectively we have

@@

\begin{theorem}[Granger Causality Equivalences]
  Given W.S.S. vector valued process $x(t)$ the following are equivalent:

  \begin{enumerate}
    \item{$x_j \gc x_i$}
    \item{$\forall \tau \in \N_+\ B_{ij}(\tau) = 0$}
    \item{$\mathcal{H}_t^{(i)} \perp \mathcal{H}_{t - 1}^{(j)}\ |\ \mathcal{H}_{t - 1}^{(i)}$}
    \item{$\mathcal{F}_{j \rightarrow i} \defeq \ln\frac{\xi_i}{\xi_{ij}} > 0$}
    \item{\hl{Wold $A(\tau)$ condition}}
    \item{Tie in the 0s of $S(\omega)^{-1}$, though this isn't immediately directed.}
  \end{enumerate}
\end{theorem}

The following propositions justify various modifications of $x(t)$
applied in practice to massage $x(t)$ into a form amenable to more
standard tools.

\begin{theorem}[Invariance Under Invertible and Deterministic Modifications]
  Let $F(z), G(z)$ describe univariate and invertible
  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
  predictable processes.  Then,

  \begin{equation}
    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
  \end{equation}
\end{theorem}

\begin{theorem}[General Invariance]
  \hl{Under what conditions is this actually true?}

  Let $\zeta(t)$ be a stationary discrete time stochastic process.  Let
  $F, G$ be (possibly nonlinear and time varying) invertible filtering
  operations and $f(t)$, $g(t)$ be perfectly predictable and such that
  $x_j(t) \defeq F(\zeta_j - f)(t), x_i(t) \defeq G(\zeta_i - g)(t)$ are W.S.S.  Then,

\begin{equation}
    x_j \gc x_i \iff \zeta_j \te \zeta_i
  \end{equation}
\end{theorem}

\begin{theorem}[Tree Recovery]
  \label{thm:tree_recovery}
  If the Granger-Causality graph $\mathcal{G}$ for $x(t)$ is a
  directed forest, then $\mathcal{G}$ can be inferred from pairwise
  causality tests alone.
\end{theorem}

\section{Structure Learning}
\subsection{Edge-Wise Grouped LASSO}
\subsection{Bayesian Posterior Thresholding}
Consider the Bayesian model

\begin{equation}
  \begin{aligned}
    (x(t)\ |\ B, \sigma_v^2, \{x(t - \tau)\}_{\tau = 1}^p) &\sim \mathcal{N}(\sum_{\tau = 1}^pB(\tau)x(t - \tau), \sigma_v^2)\\
    (B_{ij}(\tau)\ |\ G_{ij}(\tau)) &\sim G_{ij}(\tau)\mathcal{N}(0, \sigma_\beta^2) + (1 - G_{ij}(\tau))\delta_0(B_{ij}(\tau))\\
    (G_{ij}(\tau)) &\sim \mathsf{BER}(p_{ij}(\tau))\\
    \sigma_v^2 &\sim \Gamma(a_v, b_v)\\
    \sigma_\beta^2 &\sim \Gamma(a_\beta, b_\beta)
  \end{aligned}
\end{equation}

Similar models have been studied by various authors in the context of the linear regression model \hl{[(cite them)]} where it is referred to as \textit{stochastic variable selection}, or referred to as a model for Bayesian variable selection.

Since we have in mind applications where $n$ is large, sampling posterior probabilities can be prohibitively burdensome, so we can instead fit the mean-field variational approximation $p(B, G, \sigma\ |\ X) \approx q_Gq_Bq_\sigma$ and subsequently apply a thresholding operation to the posterior edge inclusion probabilities under $q_G$.

\subsection{Pairwise Minimum Error Spanning Trees}
Inspired by theorem \ref{thm:tree_recovery}, we propose the following
structure learning heuristic.

%% This algorithm should perform coordinate descent or jointly refit all the trees.

\begin{enumerate}
  \item{Set $k = 0$ and $x^0(t) = x(t)$}
  \item{Compute all pairwise causality measures $\Xi \defeq \big[\ln \frac{\xi_i}{\xi_{ij}} \big]_{i, j}$}
  \item{Find the maximum spanning arborescence\footnote{An ``arborescence'' is a french word for a tree diagram, and refers to a \textit{directed} tree in graph theory} $\mathcal{T}_k$ of a graph having edge weights $\Xi_{ij}$.}
  \item{Fit a vector LTI filter $F_k$ having graph defined by $\mathcal{T}_k$ to $x^{k - 1}(t)$ and set $x^k(t) = x^{k - 1}(t) - \hat{x}^{k - 1}(t)$}
  \item{Repeat over $k$ until a stopping criteria is met}
\end{enumerate}

\section{Application}
\section{Conclusion}
